---
title: "Crystal"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

Load the libraries
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(rgl)
knit_hooks$set(webgl = hook_webgl)
library(geometry)
library(dplyr)
library(ggplot2)
library(plotly)
RUBY = "#A01D26" #I like this color
```

Read the lattice points and correlations
```{r}
dat = read.table("~/Code/R/crystal_lattice/correlations3_chi1.4_def-3.0_famp0.08.lst")
names(dat) = c("x", "y", "z", "cor")
```

Check the dimensions
```{r}
dim(dat)
```
Separate the coordinates from the correlation
```{r}
p   = as.matrix(dat[,1:3])
cor = dat[,4]
```

Plot the histogram using `ggplot` for plotting and plotly for turning it into an interactive D3.js javascript plot 
```{r, plotly=TRUE}
gr = ggplot(dat, aes(x = cor)) +
  geom_histogram(bins=128, fill=RUBY, col=RUBY) +
  #geom_density(col = "gray10", lwd=1) +
  geom_vline(xintercept = 0.07, lty = 2, col="black", lwd=1) +
  geom_vline(xintercept = 0.143, lty = 2, col="black", lwd=1) +
  # geom_hline(yintercept = 0, col="black", lwd=0.5) +
  annotate("text", 0.077, -30, label = "0.07") +
  annotate("text", 0.152, -30, label = "0.143") +
  xlab("correlation") +
  theme_bw()
gr
# ggplotly(gr)
```

```{r, echo=FALSE}
cor_thr = 0.07
```

Observing the histogram it becomes clear that the distribution of correlations is bimodal and the split between the two modes is at around `r cor_thr`. Out of the total of `r length(cor)` unit cells there are `r sum(cor < cor_thr)` to the left of `r cor_thr` and `r sum(cor >= cor_thr)` to the right. We consider the left part as background and the right part as the crystal.

Let us plot the unit cells belonging to the crystal, using `rgl`
```{r testgl, webgl = TRUE}
p_xs = p[cor > cor_thr,]
points3d(p_xs)
```

Let's calculate the convex hull, using package `geometry`, and plot that with `rgl`
```{r, webgl = TRUE}
ts.surf <- t(convhulln(p_xs))
hullid = triangles3d(p_xs[ts.surf,1], 
                       p_xs[ts.surf,2],
                       p_xs[ts.surf,3],
                       col="green", alpha=1)
```

That was one way to segment the crystal. Another way could be to just use the unit cells themselves. Let's draw out all of the unitcells, considering that these lattice points represent the middlepoints of unitcells. We also know the unit vectors. Let's first read them in

!!! evaluation currently disabled because it takes a while to render !!!
```{r, webgl=TRUE, echo=TRUE, eval=FALSE}
#Read the 3 vectors as a 3x3 matrix. Columns are vectors
A = as.matrix(read.table("~/Code/R/crystal_lattice/realfitvectors_lat_new.lst", nrows = 3))

#Create the default cube with corners at -1 and +1
c3d = cube3d()

#Divide by 2 to make the sides unit length
c3d$vb[1:3,] = c3d$vb[1:3,]/2

#Multiply the coordinates by the unit vectors to transform the cubes into general parallelepiped, described by the unit vectors and which represent the unit cells
c3d$vb[1:3,] = A %*% c3d$vb[1:3,]

#Create an empty list to hold the cubes that are about to get drawn
cubes = vector("list", nrow(p_xs))

#Let's fill the cubes list
for(i in 1:nrow(p_xs)){
  newc = c3d #Create new copy of the cube
  newc$vb[1:3,] = newc$vb[1:3,] + p_xs[i,] #Translate the copy to the next lattice point
  cubes[[i]] = newc #Add the translated cube to the list
}

open3d()
shapelist3d(cubes, col = RUBY, alpha=0.05)
```

Looks freakin' awesome, if you ask me

```{r, webgl=TRUE, echo=TRUE, eval=FALSE}
#Create list of points that are above high threshold
p_thr = p[cor > 0.143, ]

#Create an empty list to hold the cubes that are above threshold
cubes_thr = vector("list", nrow(p_thr))

#Let's fill the cubes list
for(i in 1:nrow(p_thr)){
  newc = c3d #Create new copy of the cube
  newc$vb[1:3,] = newc$vb[1:3,] + p_thr[i,] #Translate the copy to the next lattice point
  cubes_thr[[i]] = newc #Add the translated cube to the list
}

open3d()
shapelist3d(cubes_thr, col = RUBY, alpha=1)
```


```{r, eval=FALSE}
#vector representing a plane 
a = A[,1]

rgl.triangles(p_xs[ts.surf[,4:5], 1], p_xs[ts.surf[,4:5], 2], p_xs[ts.surf[,4:5], 3])
plot3d(p_xs[ts.surf,1], p_xs[ts.surf,2], p_xs[ts.surf,3], col="green", alpha=1)
planes3d(a[1], a[2], a[3], d = d)

mass_center = c(mean(p_xs[ts.surf,1]), mean(p_xs[ts.surf,2]), mean(p_xs[ts.surf,3]))
points3d(p_xs[ts.surf,1] - mass_center[1], 
       p_xs[ts.surf,2] - mass_center[2], 
       p_xs[ts.surf,3] - mass_center[3], col="green", alpha=1)

a 
x = p_xs[ts.surf[,1]]
d = - a %*% x



      
sweep(ts_tmp, 1, c(3000, 300, 300))
```

